### 2023-12-25 sections/4-results.tex
rearrange fig 4
edits in results

Relative to fashion-mnist, glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}. 
Percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
The 95th percentile of LFD is less than 4 for all depths, and the median is less than 2 for all depths. 
Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths. 
The median hovers near 1.5 for all depths.

@@

In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift. 
We observe that the LFD starts as high as 20 (for all data) at depth 0. As shown by the needlepoint shape of the plot, the LFD for all clusters 
is about the same from depth 0 through depth 5. 
After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth. 
The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
The LFD of approximately 20 at depth 0 is what we expect for a random dataset.
To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum pairwise distance between any two points is approximately equal to maximum pairwise distance. 
As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point. 
A consequence of this is that with high probability, a point in the root cluster is a distance greater than half the root radius from the center; $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r) = |X|$, where $X$ is the dataset. 
Given our definition of LFD in Equation~\ref{eq:methods:lfd-simplified}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.

Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently show LFD.
The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 130. 
The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 130.

Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD. 
Notably however, for the frist approximately 15 depths, the 95th percentile and median LFD are very close to 0. 
Both percentiles increase sharply after depth 20, nearly reaching 16. 
LFD remains high even as depth increases, with the 95th percentile fluctuating between about 6 and 15 for depths about 20-40. 
It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60. 
The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.

\begin{figure}
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
    \subcaption{Fashion-mnist}
    \label{fig:results:fashion-mnist-lfd}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
    \subcaption{Glove-25}
    \label{fig:results:glove-25-lfd}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
    \subcaption{Sift}
    \label{fig:results:sift-lfd}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
    \subcaption{A random dataset}
    \label{fig:results:random-lfd}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
    \subcaption{Silva 18S metagenomics}
    \label{fig:results:silva-lfd}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
    \subcaption{RadioML}
    \label{fig:results:radioml-lfd}
    \end{subfigure}%

    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of LFD and weighted by the cardinalities of the clusters.}
    \label{fig:results:lfd-plots}
\end{figure}
%% end LFD plots

@@

Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance. 
Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
The horizontal axis in each figure denotes the multiplier by which we increased the cardinality of the original dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}). 
The vertical axis denotes throughput in queries per second. 
Both axes are on a logarithmic scale. 
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in our Supplement section. 
In these plots, we label the recall at each scale for algorithms which do not exhibit perfect recall; if no such labels are given for an algorithm, then that algorithm exhibits perfect recall at all scales.

Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier (i.e., scale) for the fashion-mnist, glove-25, sift, and random datasets respectively. When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider
only three decimal places.

In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the augmentation multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange. 
We also observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat (plotted in brown) for all cardinality multipliers and outperforms FAISS-IVF (plotted in pink) for all multipliers greater than $10^{5.5}$. 
Though state-of-the-art algorithms HNSW (plotted in gray), and Annoy (plotted in purple) have higher throughput than CAKES for all multipliers, we note that CAKES exhibits perfect recall for all multipliers, while hnsw and annoy exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}. 
For example, at a cardinality multiplier as low as eight, hnsw and annoy have recall of 0.525 and 0.857 respectively, while CAKES exhibits perfect recall.
Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance for nearly all cardinality multipliers.

With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high cardinality multipliers. 
Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
We also see that while HNSW and Annoy have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multupliers, as shown in Table~\ref{table:results:ann-glove-25}.
In particular, at a cardinality multiplier as low as two, HNSW and Annoy have recall as low as 0.607 and 0.832 respectively, while CAKES exhibits perfect recall.
On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN, plotted in green, is consistently the best-performing CAKES algorithm.

In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low cardinality multipliers, our algorithms begin outperforming linear search at a multiplier of approximately 7. 
Again, we observe that CAKES outperforms state-of-the-art similarity search algorithm FAISS-Flat for all cardinality multipliers and outperforms FAISS-IVF for all multipliers greater than $10^{7}$. 
For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than 7. 
Similar to the results with fashion-mnist and glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES for all multipliers, CAKES has better recall. 
Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and Annoy have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.

With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibt the worst throughput out of all algorithms, at all cardinality multipliers. 
As with the real datasets, HNSW and Annoy exhibit the highest throughput of any of the algorithms. 
As with the real datasets, CAKES exhibits perfect recall at all cardinality multipliers. 
HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
in particular, with a cardinality multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.

@@

\begin{figure}
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
\subcaption{Fashion-mnist for $k=10$.}
\label{fig:results:fashion-mnist-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
\subcaption{Glove-25 for $k=10$.  }
\label{fig:results:glove-25-scaling}
\end{subfigure}%
\vspace{1em}
\\
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
\subcaption{Sift for $k=10$.    }
\label{fig:results:sift-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
\subcaption{A random dataset for $k=10$.
}
\label{fig:results:random-scaling}
\end{subfigure}%
\vspace{1em}
\\
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
\subcaption{Silva for $k=10$.    }
\label{fig:results:silva-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
\subcaption{RadioML for $k=10$ at SnR = 10dB.
}
\label{fig:results:radioml-scaling}
\end{subfigure}%

>>>>

Relative to fashion-mnist, glove-25 has low LFD, as shown in Figure~\ref{fig:results:glove-25-lfd}.
Percentile lines for glove-25 are flatter and lower, indicating that the LFD is lower across the entire dataset, and that the LFD does not vary as much by depth.
The 95th percentile of LFD is less than 3 for all depths, and the median is less than 2 for all depths. 
Before depth 25, the 95th percentile hovers near 2, and from depth 20 onward, it hovers near 3.5 before dipping sharply at the maximum depths.
The median hovers near 1.5 for all depths.

@@

In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift.
We observe that the LFD starts as high as 20 (for all data) at depth 0.
As shown by the needlepoint shape of the plot, the LFD for all clusters is about the same from depth 0 through depth 5.
After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth.
The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
The LFD of approximately 20 at depth 0 is what we expect for this random dataset.
To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum and maximum pairwise distances between any two points are approximately equal.
As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point.
A consequence of this is that with high probability, a point in the root cluster is a distance greater than half the root radius from the center; $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r) = |X|$, where $X$ is the dataset.
Given our definition of LFD in Equation~\ref{eq:methods:lfd-simplified}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.

Silva, as shown in Figure~\ref{fig:results:silva-lfd}, exhibits consistently low LFD.
The 95th percentile is less than about 3 for all depths, hovering near 1 for depths 40 through 128.
The median reaches its peak at 2 for about depth 10 and remains close to 0 for depths 40 through 128.

Relative to the other datasets, RadioML, as displayed in Figure~\ref{fig:results:radioml-lfd}, exhibits high LFD.
Notably however, for the first approximately 15 depths, the 95th percentile and median LFD are very close to 0.
Both percentiles increase sharply after depth 20, nearly reaching 16.
LFD remains high even as depth increases, with the 95th percentile fluctuating between about 6 and 15 for depths about 20-40.
It then spikes up again to about 10 at around depth 45 before decreasing linearly to 1 at depth 60.
The median LFD follows a similar pattern of peaks and troughs, fluctuating between about 3 and 15 for depths 20-40, spiking at about depth 50 to 5, and then decreasing approximately linearly to 1 at depth 60.

\begin{figure}
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/fashion-mnist-60000-lfd.png}\\
    \subcaption{Fashion-mnist}
    \label{fig:results:fashion-mnist-lfd}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/glove-25-1183514-lfd.png}\\
    \subcaption{Glove-25}
    \label{fig:results:glove-25-lfd}
    \end{subfigure}
    \vspace{1em}
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/sift-1000000-lfd.png}\\
    \subcaption{Sift}
    \label{fig:results:sift-lfd}
    \end{subfigure}%
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/radio-ml-97920-lfd.png}\\
    \subcaption{RadioML}
    \label{fig:results:radioml-lfd}
    \end{subfigure}%
    \\
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/silva-2224640-lfd.png}\\
    \subcaption{Silva 18S}
    \label{fig:results:silva-lfd}
    \end{subfigure}%  
    \begin{subfigure}[b]{0.47\textwidth}
    \includegraphics[width=0.95\textwidth]{images/lfd_plots/random-1000000-lfd.png}\\
    \subcaption{A random dataset}
    \label{fig:results:random-lfd}
    \end{subfigure}
    \vspace{1em}
    \caption{Local fractal dimension vs. cluster depth across six datasets, grouped by decile of local fractal dimension and weighted by the cardinalities of the clusters.
    The last dataset is randomly generated.
    {\color{red} TODO: Remove the headers from the plots themselves and rely on subcaptions instead.}}
    \label{fig:results:lfd-plots}
\end{figure}

@@

Figures~\ref{fig:results:fashion-mnist-scaling}, ~\ref{fig:results:glove-25-scaling}, and ~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of the following ann-benchmark datasets: fashion-mnist under Euclidean distance, glove-25 under cosine distance, and sift under Euclidean distance.
Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
The horizontal axis in each figure shows the cardinality of the augmented dataset with synthetic points (see Section \ref{subsec:methods:synthetic-data}).
The left-most point on each line is on the original dataset without any synthetic augmentation.
The vertical axis denotes throughput in queries per second.
Both axes are on a logarithmic scale.
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k=100$ can be found in the Supplement.
In these plots, we label the recall for algorithms which do not exhibit perfect recall;
if no such labels are given for an algorithm, then that algorithm exhibits perfect recall.

Tables \ref{table:results:ann-fashion}, \ref{table:results:ann-glove-25}, \ref{table:results:ann-sift}, and \ref{table:results:ann-random} show the throughput and recall of CAKES's algorithms at each cardinality multiplier for the fashion-mnist, glove-25, sift, and random datasets respectively.
When reporting recall, we use 1.000* to denote that the recall is imperfect, but rounds to 1.000 when we consider only three decimal places.

In Figure~\ref{fig:results:fashion-mnist-scaling}, which shows results for fashion-mnist, we observe that as the multiplier increases, the CAKES algorithms (GreedySieve in blue, RepeatedRnn in green, and Sieve in red) exhibit higher throughput than na\"{i}ve linear search, plotted in orange.
We also observe that CAKES outperforms FAISS-IVF (plotted in pink) for all multipliers greater than one.
Though state-of-the-art algorithms HNSW (plotted in gray), and ANNOY (plotted in purple) have higher throughput than CAKES for all cardinalities, we note that CAKES always exhibits perfect recall while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{table:results:ann-fashion}.
For example, at a multiplier as low as eight, HNSW and ANNOY have recall of 0.525 and 0.857 respectively,
and this recall gets worse as the multiplier increases.
Meanwhile, while CAKES exhibits perfect recall.
Amongst the CAKES algorithms, GreedySieve consistently performs best on this dataset, exhibiting sublinear time performance with increasing multipliers.

With glove-25 (Figure~\ref{fig:results:glove-25-scaling}), we observe that each of CAKES' algorithms outperforms linear search at high multipliers.
Similar to the results for fashion-mnist, we see that CAKES outperforms FAISS-IVF for all multipliers greater than four.
We see once again that HNSW and ANNOY have higher throughput than CAKES for all multipliers, CAKES exhibits better recall at all multipliers, as shown in Table~\ref{table:results:ann-glove-25}.
In particular, at a multiplier as low as two, HNSW and ANNOY have recall as low as 0.607 and 0.832 respectively, and this recall continues to decrease as we increase the multiplier.
In contrast, CAKES exhibits perfect recall at all multipliers.
On this dataset, unlike with fashion-mnist, Repeated $\rho$-NN, plotted in green, is consistently the best-performing CAKES algorithm.

In Figure~\ref{fig:results:sift-scaling}, which displays the scaling behavior of algorithms on the sift dataset, we observe that while linear search performs best for low multipliers, our algorithms begin to outperform linear search at a multiplier of eight.
Again, we observe that CAKES outperforms FAISS-IVF for all multipliers greater than four. 
For this dataset, Sieve, plotted in red, is the highest throughput CAKES algorithm for all multipliers greater than right.
Similar to the results with fashion-mnist and glove-25, we observe that while HNSW and Annoy exhibit higher throughput than CAKES for all multipliers, CAKES has better recall.
Table~\ref{table:results:ann-sift} shows that even with a multiplier of 1 (i.e., the original dataset), HNSW and Annoy have recall as low as 0.782 and 0.686 respectively, while CAKES exhibits perfect recall.

With the random dataset, as seen in Figure~\ref{fig:results:random-scaling}, we observe that the three CAKES algorithms exhibit the worst throughput out of all algorithms, at all multipliers. 
As with the real datasets, HNSW and ANNOY exhibit the highest throughput of any of the algorithms and CAKES exhibits perfect recall at all multipliers. 
HNSW and Annoy exhibit \emph{much} lower recall on this random dataset than on the real datasets; 
in particular, with a multiplier of 1, HNSW and Annoy have recall as low as 0.060 and 0.028 respectively, as reported in Table~\ref{table:results:ann-random}.

@@

\begin{figure}
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/fashion-mnist-knn-10.png}
\subcaption{Fashion-mnist for $k=10$.}
\label{fig:results:fashion-mnist-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/glove-25-knn-10.png}
\subcaption{Glove-25 for $k=10$.  }
\label{fig:results:glove-25-scaling}
\end{subfigure}%
\vspace{1em}
\\
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/sift-knn-10.png}
\subcaption{Sift for $k=10$.    }
\label{fig:results:sift-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/radio-ml-knn-10.png}
\subcaption{RadioML for $k=10$ at SnR = 10dB.
}
\label{fig:results:radioml-scaling}
\end{subfigure}%
\vspace{1em}
\\
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/silva-knn-10.png}
\subcaption{Silva for $k=10$.    }
\label{fig:results:silva-scaling}
\end{subfigure}%
\begin{subfigure}[b]{0.47\textwidth}
\includegraphics[width=0.95\textwidth]{plots/random-knn-10.png}
\subcaption{A random dataset for $k=10$.
}
\label{fig:results:random-scaling}
\end{subfigure}%
